---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rentrezaddon

__Warning: This package is still under development !__

A package comprising functions to execute [standalone Blast+](https://www.ncbi.nlm.nih.gov/books/NBK279690/)
searches from within R (though executive and search path have to be set up separately),
to aid workflows based on [rentrez](https://docs.ropensci.org/rentrez/) and 
[Biostrings](https://bioconductor.org/packages/release/bioc/html/Biostrings.html).


## Installation

You can install the released version of rentrezaddon from GitHub.

```{r installation, eval = FALSE}
# should not work, since the repo is private
# devtools::install_github("mirkko-hub/rentrezaddon")
```

```{r}
library(tidyverse, warn.conflicts = FALSE)
library(magrittr, warn.conflicts = FALSE)
library(rentrezaddon)
```

## Examples

#### Blast searches - Prerequisites

* local installation of current release of [standalone Blast+](https://www.ncbi.nlm.nih.gov/books/NBK279690/)
* databases generated using 'makeblastdb' (add '-parse_seqids' to work with blast_cmd)
* some query file in fasta format
* search path to executives are given in the body of the function, which have to 
  be modified accordingly.
  
_Proper setup should enable something like this to work_

(files are local only and not provided here, since everyone else would be lacking the BLAST+ executive anyways!)

```{r blast}
# databases to search:
nuccores <- readr::read_csv("~/Documents/NCBI_thesis_2/analysis/20201120_assembly_information_rbcl_form.csv") %$% caption
db_path <- paste0("~/Documents/NCBI_thesis_2/nuccore/db/", nuccores)[1:100]

# query
ccmm <- "~/Documents/NCBI_thesis_2/queries/pcc7120_ccmm.faa"

# search
blastn_ccmm <- purrr::map_dfr(db_path, query_blast, query = ccmm, blast = tblastn)
head(blastn_ccmm)

# works with conserved domain searches, too
blast_cdd <- query_blast("/home/mirkko/Documents/NCBI/ncbi_databases/cdd/Cdd_NCBI", query = ccmm, blast = blastp)
head(blast_cdd)

### check ccds with query_blastdbcmd_cdd -- see below!
```

```{r blastdbcmd}
# works only with indexed databases
# to generate an indexed daabase, you have to include '-parse_seqids' in 'makeblastdb' command!
query_blastdbcmd(db = "/home/mirkko/Documents/r_packages/test_data/example_db",
                 entry = "sequence_1",
                 dbtype = "nucl",
                 filepath = "data-raw/example_blastdbcmd", range = "1-10", strand = "minus")
Biostrings::readDNAStringSet("data-raw/example_blastdbcmd")

# if no indexed database available / you don't want to generate it / above function does not work fo you
# -> try 'blast_to_xstringset()'
```
```{r blast_to_xstringset}
# get sequences (xstringset) from identified ranges (blast)
# generate example_xstringset and example_blast:
example_xstringset <- system.file("extdata", "example_xstringset.fna" , package = "rentrezaddon", mustWork = TRUE) %>% Biostrings::readDNAStringSet(.)
example_xstringset
example_blast

# if DNA, output is ALWAYS in 5'-3' direction no matter which DNA strand the region is on!
example_blast %>%
blast_to_xstringset(., xstringset = example_xstringset)
```


```{r query_blastdbcmd_cdd}
# get further info to cdd motifs
blast_cdd$sseqid %>% purrr::map_dfr(., query_blastdbcmd_cdd, db = "/home/mirkko/Documents/NCBI/ncbi_databases/cdd/Cdd_NCBI")
```

#### Some more Biostrings helpers

* _have a fasta/uniprot/genbank/etc. header?_
* __just wanna rename the files with id and remove that other crap?__

why don't you try this?
```{r}
example_header <- system.file("extdata", "example_headers.faa" , package = "rentrezaddon", mustWork = TRUE) %>% Biostrings::readAAStringSet(.)
example_header %>% names()
example_header %>% names() %>% map_dfr(., extract_header_2)
example_header %>% `names<-`(names(.) %>% map_dfr(., extract_header_2) %$% faa_id)
```
#### Some more rentrez help?

```{r}
esummary_2 <- rentrez::entrez_summary(db = "protein", id = c("119370761", "73919686"))
esummary_2 %>% purrr::map_dfr(., extract_esummary_protein)
```

